\ \\ \LARGE\bf Generating Lambda Term Individuals in Typed Genetic Programming Using Forgetful A*

### 2014-01-19 tex/cec14/article.tex
cecky

We made tree experiments comparing performance of standard 
\textit{ramped half-and-half strategy} with our
\textit{geometric strategy} (with default parameter $q=0.75$). 
Each experiment consisted of 50 independent runs 
of GP algorithm. Each run had maximally 51 generations and population size
was 500 individuals.
In those experiments we analyze the ability of the system to produce 
a correct solution, computational cost estimates and average size 
of an individual throughout generations. 
For the first two metrics we use the popular measurement  
methods within GP field --- the \textit{performance curves}
described in \cite{koza92} --- $P(M,i)$ (cumulative probability of success) 
and $I(M,i,z)$ (the total number of individuals that must be processed 
to yield a correct solution with probability $z =99\%$).

@@

\begin{figure}[h!]
  \caption{Performance curves for Simple symbolic regression.}
  \centering
    \includegraphics[scale=0.1]{imgs/SSR_FINAL.png}
\end{figure}
Fitness function is computed as follows.
$$ fitness(f) =  \sum\limits_{i=1}^{20}{ \vert f(x_i)-y_i }\vert   $$
where $(x_i,y_i)$ are 20 data samples from $[-1,1]$, such that $y_i = f_t(x_i)$.
An individual $f$ such that $\vert f(x_i)-y_i \vert < 0.01 $ for all data samples is 
considered as a correct individual.

@@

\begin{figure}[h!]
  \caption{Performance curves for Artificial ant problem.}
  \centering
    \includegraphics[scale=0.1]{imgs/ANT_FINAL.png}
\end{figure}

@@

@@

Previous two problems were instances of standard GP, i.e. their building symbols
obeyed the closure requirement. This third experiment will break the closure requirement, thus ideal for testing typed GP techniques. 
The even-parity function is a boolean function taking as inputs $N$
boolean values and returning \textit{True} if an even number of inputs 
are \textit{True}. For odd number it returns \textit{False}.
This problem has been used by many researchers
as a benchmark for GP. We compare our results with that obtained by Yu in \cite{yu01},
where is presented approach to evolve recursive and modular programs
by use of higher-order functions and $\lambda$-abstractions.
We use very similar set of \textit{building symbols} as in \cite{yu01}. 
\begin{align*}
\tau = [Boo&l] \ar Bool\\
\Gamma = \{
  and   &: Bool \ar Bool \ar Bool                              ,
  or    : Bool \ar Bool \ar Bool                              ,\\
  nand  &: Bool \ar Bool \ar Bool                              ,
  nor    : Bool \ar Bool \ar Bool                              ,\\
  foldr &: (Bool \ar Bool \ar Bool) \ar Bool \ar [Bool] \ar Bool ,\\
  head' &: [Bool] \ar Bool                                   ,
  tail' : [Bool] \ar [Bool]                              \}
\end{align*}
\begin{figure}[h!]
  \caption{Performance curves for Even parity problem.}
  \centering
    \includegraphics[scale=0.1]{imgs/EP_FINAL.png}
\end{figure}

>>>>

We made tree experiments comparing performance of standard 
\textit{ramped half-and-half strategy} with our
\textit{geometric strategy} (with default parameter $q=0.75$). 
Each experiment consisted of 50 independent runs 
of GP algorithm. Each run had maximally 51 generations and population size
was 500 individuals.
In those experiments we analyze the fitness of the best individual,
the average size of an individual,
the ability of the system to produce a correct solution 
and computational cost estimates throughout generations. 
For the last two metrics we use the popular measurement  
methods within GP field --- the \textit{performance curves}
described in \cite{koza92} --- $P(M,i)$ (cumulative probability of success) 
and $I(M,i,z)$ (the total number of individuals that must be processed 
to yield a correct solution with probability $z =99\%$).
Difference of those two generating methods 
is statistically analyzed by the Welch t-test
by comparing the fitness values of the best individuals of the run. 
Graphs showing mean values of the fitness of the best individual and
the average size of an individual throughout generations
contain error bars representing the standard error of the mean (SEM).

@@

\begin{figure}[h!]
  \caption{Graphs for Simple symbolic regression.}
  \centering
    \includegraphics[scale=0.15]{imgs/cec/SSR.PNG}
\end{figure}

~\\Fitness function is computed as follows.

$$ fitness(f) =  \frac{1}{1+ \sum\limits_{i=1}^{20}{ \vert f(x_i)-y_i }\vert }  $$
where $(x_i,y_i)$ are 20 data samples from $[-1,1]$, such that $y_i = f_t(x_i)$.
An individual $f$ such that $\vert f(x_i)-y_i \vert < 0.01 $ for all data samples is 
considered as a correct individual.

@@

\begin{figure}[h!]
  \caption{Graphs for Artificial ant problem.}
  \centering
    \includegraphics[scale=0.15]{imgs/cec/ANT.PNG}
\end{figure}

@@

\begin{table}[t]
\centering
\begin{tabular}{|l|cc|}
\hline
& Ramped half-and-half & Geometric \\
\hline
Mean & 0.7999030182 & 0.8174067512 \\
SD	 & 0.1521369287	& 0.1826109800 \\
SEM	 & 0.0215154108	& 0.0258250925 \\
N	 & 50 & 50  \\
\hline
t-value & 0.5207 & \\
p-value & 0.6038 & \\
\hline
\end{tabular}
\caption{Statistical analysis of Simple symbolic regression.
By conventional criteria, this difference is considered to be not statistically significant.}
\end{table}

\begin{table}[t]
\centering
\begin{tabular}{|l|cc|}
\hline
& Ramped half-and-half & Geometric \\
\hline
Mean & 66.20 & 79.78 \\
SD	 & 12.33 & 8.77  \\
SEM	 & 1.74	 & 1.24  \\
N	 & 50    & 50    \\
\hline
t-value &  6.3479           & \\
p-value &  less than 0.0001 &  \\
\hline
\end{tabular}
\caption{Statistical analysis of Artificial ant problem.
 By conventional criteria, this difference is considered to be extremely statistically significant. }
\end{table}

\begin{table}[t]
\centering
\begin{tabular}{|l|cc|}
\hline
& Ramped half-and-half & Geometric \\
\hline
Mean & 0.895000	& 0.930000 \\
SD	 & 0.075170	& 0.078490 \\
SEM	 & 0.010631	& 0.011100 \\
N	 & 50    & 50    \\
\hline
t-value &  2.2772 & \\
p-value &  0.0250 &  \\
\hline
\end{tabular}
\caption{Statistical analysis of Even parity problem. 
By conventional criteria, this difference is considered to be statistically significant.}
\end{table}

@@

Previous two problems were instances of standard GP, i.e. their building symbols
obeyed the closure requirement. This third experiment will break the closure requirement, thus ideal for testing typed GP techniques. 
The even-parity function is a boolean function taking as inputs $N$
boolean values and returning \textit{True} if an even number of inputs 
are \textit{True}. For odd number it returns \textit{False}.
This problem has been used by many researchers
as a benchmark for GP. We compare our results with that obtained by Yu in \cite{yu01},
where is presented approach to evolve recursive and modular programs
by use of higher-order functions and $\lambda$-abstractions.
We use very similar set of \textit{building symbols} as in \cite{yu01}. 
\begin{align*}
\tau = [Boo&l] \ar Bool\\
\Gamma = \{
  and   &: Bool \ar Bool \ar Bool                              ,\\
  or    &: Bool \ar Bool \ar Bool                              ,\\
  nand  &: Bool \ar Bool \ar Bool                              ,\\
  nor   &: Bool \ar Bool \ar Bool                              ,\\
  foldr &: (Bool \ar Bool \ar Bool) \\
        &~~~\ar Bool \ar [Bool] \ar Bool ,\\
  head' &: [Bool] \ar Bool                                   ,\\
  tail' &: [Bool] \ar [Bool]                              \}
\end{align*}
\begin{figure}[h!]
  \caption{Graphs for Even parity problem.}
  \centering
    \includegraphics[scale=0.15]{imgs/cec/EP.PNG}
\end{figure}
